---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## acton

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/cyipt/acton.svg?branch=master)](https://travis-ci.com/cyipt/acton)
[![CircleCI](https://circleci.com/gh/cyipt/acton.svg?style=svg)](https://circleci.com/gh/cyipt/acton)
<!-- badges: end -->

## Introduction

**ACTON** stands for active transport options for new developments.
ACTON is a research project to provide evidence for local authorities, developers and civil society groups to support planning and investment in sustainable transport infrastructure in and around new developments.
To make the results of the research more reproducible and accessible to others, it is also an R package.

In first place, ACTON is about evidence-based planning and that means it is about data.
ACTON provides easy access to actionable data from 4 main sources, each of which is already in the public domain, which had never before been made available in a single place or analysed together to support planning, as illustrated in the schematic diagram below.


```{r workflow, echo=FALSE}
nomnoml::nomnoml("
#direction: down
#.box: fill=#8f8 visual=ellipse
[PlanIt]-[<box>ACTON]
[Accessibility]-[ACTON]
[Travel/Demographics]-[ACTON]
[Route]-[ACTON]
")
```

The four main data sources shown above are:

- PlanIt data on new developments based on planning documents submitted to local government planning departments
- Accessibility data, including travel times by mode for different trip purposes, e.g. the average time taken to cycle to the shops, for small administrative zones (MSOA and LSOA level)
- Trave/demographic data from surveys (the 2011 Census in the first instance), providing information about the number of people in different groups in administrative zones and likely work destinations at the origin-destination level (plus data derived from the Census via the Propensity to Cycle Tool project)
- Route data from routing services such as CycleStreets.net, providing information on the transport networks surrounding new development sites, with variables such as busyness and speed limits along roads nearby new and planned development sites

ACTON seeks to make these datasets more widely accessible and actionable by combining them and using them to assess walking and cycling provision in and around new developments, to inform policies and investment

For more information about the research project, see the in-progress [ACTON report](https://cyipt.github.io/acton/articles/the-acton-project.html) and [case study](https://cyipt.github.io/acton/articles/case-studies.html) articles.

## Installing the R package

To install the `acton` package, run the following commands in an R console (see [here](https://docs.ropensci.org/stats19/articles/stats19-training-setup.html) for information on installing R):

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("cyipt/acton", dependencies = "Suggests")
```

## Setup instructions

To get routes from CycleStreets.net, you will need to set-up an API key called CYCLESTREETS with `usethis::edit_r_environ()`, as documented here: https://docs.ropensci.org/stplanr/reference/route_cyclestreets.html#details

## Brief demo

The package can be used to get data on new developments as follows:

```{r}
library(acton)
# data from specific postcode
planning_data = get_planit_data(pcode = "LS2 9JT", limit = 2, app_size = "large")
planning_data
planning_data$name
planning_data$description
```

## Documentation

For a more detailed overview explaining how to use the package see the [`acton` vignette](https://cyipt.github.io/acton/articles/acton.html).

For results of research into active travel opportunities in and around new developments in case study regions, see the [`case-studies` vignette](https://cyipt.github.io/acton/articles/case-studies.html).

For documentation on each of the package's functions, see the [Reference page](https://cyipt.github.io/acton/reference/index.html).

## Citing the work

TBC.

