---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## acton

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/cyipt/acton.svg?branch=master)](https://travis-ci.com/cyipt/acton)
[![CircleCI](https://circleci.com/gh/cyipt/acton.svg?style=svg)](https://circleci.com/gh/cyipt/acton)
<!-- badges: end -->

## Introduction

**ACTON** is a research project to provide evidence for local authorities, developers and civil society groups to support planning and investment in sustainable transport infrastructure in and around new developments.
To make the results of the research more reproducible and accessible to others, we have also created an R package, which is described below.

For information about the wider research project, see the in-progress [ACTON report](https://cyipt.github.io/acton/articles/the-acton-project.html) and [case study](https://cyipt.github.io/acton/articles/case-studies.html) articles.

## Installing the R package

To install the `acton` package, run the following commands in an R console (see [here](https://docs.ropensci.org/stats19/articles/stats19-training-setup.html) for information on installing R):

```{r, eval=FALSE}
install.packages("remotes")
remotes::install_github("cyipt/acton", dependencies = "Suggests")
```

## Setup instructions

To get routes from CycleStreets.net, you will need to set-up an API key called CYCLESTREETS with `usethis::edit_r_environ()`, as documented here: https://docs.ropensci.org/stplanr/reference/route_cyclestreets.html#details

## Brief demo

The package can be used to get data on new developments as follows:

```{r}
library(acton)
# data from specific postcode
planning_data = get_planit_data(pcode = "LS2 9JT", limit = 2)
planning_data
planning_data$name
planning_data$description
```

## Documentation

For a more detailed overview explaining how to use the package see the [`acton` vignette](https://cyipt.github.io/acton/articles/acton.html).

For results of research into active travel opportunities in and around new developments in case study regions, see the [`case-studies` vignette](https://cyipt.github.io/acton/articles/case-studies.html).

For documentation on each of the package's functions, see the [Reference page](https://cyipt.github.io/acton/reference/index.html).

## Citing the work

TBC.

