---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## acton

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/cyipt/acton.svg?branch=master)](https://travis-ci.com/cyipt/acton)
[![CircleCI](https://circleci.com/gh/cyipt/acton.svg?style=svg)](https://circleci.com/gh/cyipt/acton)
<!-- badges: end -->

## Introduction

**Acton** is a research project to provide evidence for local authorities, developers and civil society groups to support planning and investment in sustainable transport infrastructure in and around new developments.
To make the results of the research more reproducible and accessible to others, we have also created an R package, which is the focus of this README.

For information about the wider research project, see the [`acton-project`](https://cyipt.github.io/acton/articles/the-acton-project.html) vignette.

For a more detailed overview explaining how to use the package see the [`acton` vignette](https://cyipt.github.io/acton/articles/acton.html).

For results of research into active travel opportunities in and around new developments in case study regions, see the [`case-studies` vignette](https://cyipt.github.io/acton/articles/case-studies.html).


## Installation

To install the `acton` package, run the following commands:

```{r, eval=FALSE}
install.packages(remotes)
remotes::install_github("cyipt/acton", dependencies = "Suggests")
```

## Setup instructions

To get routes from CycleStreets.net, you will need to set-up an API key called CYCLESTREETS with `usethis::edit_renviron()`, as documented here: https://docs.ropensci.org/stplanr/reference/route_cyclestreets.html#details

## Brief demo

The package can be used to get data on new developments as follows:

```{r}
library(acton)
# data from specific postcode
planning_data = get_planit_data(pcode = "LS2 9JT", limit = 2)
planning_data
planning_data$name
planning_data$description
```

## Citing the work

TBC.

