---
title: "Introducing the acton package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introducing the acton package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(acton)
```


# Getting planning data

To investigate the accessibility of new housing developments, it is first necessary to identify where new homes are being built. 

There are over 300 local planning authorities across England, including metropolitan boroughs, London boroughs, unitary authorities and non-metropolitan districts. Each of these planning authorities keeps an online record of current and historic planning applications, such as applications to build new homes. The data recorded will include information about the status of the application, the site location, related applications, a description of the proposed application and links to download associated files such as masterplans and transport assessments. 

However, there is no real standardisation of how these planning applications are recorded. For example, various different types of application exist, such as full, outline, reserved matters, and applications for tree works. Yet the codes that record application type vary from one council to the next.

Moreover, there is no systematic national recording of where homes are being built. The Ministry of Housing, Communities & Local Government produces statistics on the number of new build housing starts and completions in each local planning authority, but this simply records the number of homes in each district. There is no indication of exactly where these homes are being built.

To do this, the PlanIt website (https://www.planit.org.uk/) was developed. This maps current and historic planning applications across the UK, scraping the data from council websites.

# Getting accessibility data
